---
title: "Citing 'easystats'"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Citing 'easystats'}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
options(
  knitr.kable.NA = "",
  digits = 2
)

knitr::opts_chunk$set(
  comment = ">",
  dpi = 450,
  message = FALSE,
  warning = FALSE
)
```

# Citation

How to reference *easystats*?

1. **Cite specific packages**

The most parsimonious approach is to cite only the particular package that
helped you, e.g., 

> "using *bayestestR* (Makowski, Ben-Shachar, & Lüdecke, 2019)".

2. **Cite the whole ecosystem** ❤️

Want to credit the whole *easystats* ecosystem of packages? 

You can use the convenient `cite_easystats()` function:

```{r eval=FALSE}
# Cite just the 'easystats' umbrella package:
easystats::cite_easystats()

# Cite all easystats packages:
easystats::cite_easystats(packages = "all")

# Cite specific packages:
easystats::cite_easystats(packages = c("insight", "parameters", "performance"))

# For markdown/RMarkdown documents:
easystats::cite_easystats(format = "markdown")

# For LaTeX/BibLaTeX:
easystats::cite_easystats(format = "biblatex")
```

Alternatively, this can be done with a sentence like the following:

> Data processing was carried out with R (R Core Team, 2019) and the *easystats*
ecosystem (Lüdecke, Waggoner, & Makowski, 2019; Lüdecke, Ben-Shachar, Patil,
Waggoner, & Makowski, 2021; Makowski, Ben-Shachar, & Lüdecke, 2019; Makowski,
Ben-Shachar, Patil, & Lüdecke, 2020; Lüdecke, Ben-Shachar, Patil, & Makowski,
2020; Lüdecke et al., 2021; Ben-Shachar, Lüdecke, & Makowski, 2020)

- Lüdecke, D., Waggoner, P. D., & Makowski, D. (2019). insight: A Unified
  Interface to Access Information from Model Objects in R. _Journal of Open
  Source Software_, *4*, 1412. https://doi.org/10.21105/joss.01412

- Makowski, D., Ben-Shachar, M.S., & Lüdecke, D. (2019). bayestestR: Describing
  Effects and their Uncertainty, Existence and Significance within the Bayesian
  Framework. _Journal of Open Source Software_, *4*(40), 1541.
  https://doi.org/10.21105/joss.01541

- Makowski, D., Ben-Shachar, M.S., Patil, I., & Lüdecke, D. (2019). Methods and
  Algorithms for Correlation Analysis in R. _Journal of Open Source Software_,
  *5*(51), 2306. https://doi.org/10.21105/joss.02306

- Lüdecke, D., Ben-Shachar, M.S., Patil, I., & Makowski, D. (2020). parameters:
  Extracting, Computing and Exploring the Parameters of Statistical Models using
  R. _Journal of Open Source Software_, *5*(53), 2445.
  https://doi.org/10.21105/joss.02445

- Ben-Shachar, M.S., Lüdecke, D., Makowski, D. (2020). effectsize: Estimation of
  Effect Size Indices and Standardized Parameters. _Journal of Open Source
  Software_, *5*(56), 2815. https://doi.org/10.21105/joss.02815

- Lüdecke, D., Ben-Shachar, M.S., Patil, I., Waggoner, P. D., & Makowski, D.
  (2021). performance: An R Package for Assessment, Comparison and Testing of
  Statistical Models. _Journal of Open Source Software_, _6_(60), 3139.
  https://doi.org/10.21105/joss.03139

- Lüdecke, D., Patil, I., Ben-Schachar, M.S., Wiernik, B.M., Waggoner, P., &
  Makowski, D. (2021). see: An R Package for Visualizing Statistical Models.
  _Journal of Open Source Software_, _6_(64), 3393,
  https://doi.org/10.21105/joss.03393

<details> <summary>Click here to see the corresponding APA and bibtex
entries</summary>

<br>

```
@article{ludecke2019insight,
	journal = {Journal of Open Source Software},
	doi = {10.21105/joss.01412},
	issn = {2475-9066},
	number = {38},
	publisher = {The Open Journal},
	title = {insight: A Unified Interface to Access Information from Model Objects in R},
	url = {http://dx.doi.org/10.21105/joss.01412},
	volume = {4},
	author = {Lüdecke, Daniel and Waggoner, Philip and Makowski, Dominique},
	pages = {1412},
	date = {2019-06-25},
	year = {2019},
	month = {6},
	day = {25}
}


@article{makowski2019bayestestr,
	title = {{bayestestR}: {Describing} {Effects} and their {Uncertainty}, {Existence} and {Significance} within the {Bayesian} {Framework}},
	volume = {4},
	issn = {2475-9066},
	shorttitle = {{bayestestR}},
	url = {https://joss.theoj.org/papers/10.21105/joss.01541},
	doi = {10.21105/joss.01541},
	number = {40},
	urldate = {2019-08-13},
	journal = {Journal of Open Source Software},
	author = {Makowski, Dominique and Ben-Shachar, Mattan S. and Lüdecke, Daniel},
	month = aug,
	year = {2019},
	pages = {1541}
}

@article{makowski2020correlation,
  doi={10.21105/joss.02306},
  title={Methods and Algorithms for Correlation Analysis in R},
  author={Makowski, Dominique and Ben-Shachar, Mattan S. and Patil, Indrajeet and L{\"u}decke, Daniel},
  journal={Journal of Open Source Software},
  volume={5},
  number={51},
  pages={2306},
  year={2020}
}

@article{ludecke20202parameters,
  title = {parameters: Extracting, Computing and Exploring the Parameters of Statistical Models using {R}.},
  volume = {5},
  doi = {10.21105/joss.02445},
  number = {53},
  journal = {Journal of Open Source Software},
  author = {Daniel Lüdecke and Mattan S. Ben-Shachar and Indrajeet Patil and Dominique Makowski},
  year = {2020},
  pages = {2445},
}

@article{benchashar2020effectsize,
  title = {{e}ffectsize: Estimation of Effect Size Indices and Standardized Parameters},
  author = {Mattan S. Ben-Shachar and Daniel Lüdecke and Dominique Makowski},
  year = {2020},
  journal = {Journal of Open Source Software},
  volume = {5},
  number = {56},
  pages = {2815},
  publisher = {The Open Journal},
  doi = {10.21105/joss.02815},
  url = {https://doi.org/10.21105/joss.02815},
}

@Article{Lüdecke2021performance,
    title = {{performance}: An {R} Package for Assessment, Comparison and Testing of Statistical Models},
    author = {Daniel Lüdecke and Mattan S. Ben-Shachar and Indrajeet Patil and Philip Waggoner and Dominique Makowski},
    year = {2021},
    journal = {Journal of Open Source Software},
    volume = {6},
    number = {60},
    pages = {3139},
    doi = {10.21105/joss.03139},
  }
  
@Article{Lüdecke2021see,
    title = {{see}: An {R} Package for Visualizing Statistical Models},
    author = {Daniel Lüdecke and Indrajeet Patil and Mattan S. Ben-Shachar and Brenton M. Wiernik and Philip Waggoner and Dominique Makowski},
    journal = {Journal of Open Source Software},
    year = {2021},
    volume = {6},
    number = {64},
    pages = {3393},
    doi = {10.21105/joss.03393},
  }
  
```
</details>
