---
title: "easystats: A Unified Framework for Statistical Analysis in R"
tags:
  - R
authors:
- affiliation: 1
  name: Daniel Lüdecke
  orcid: 0000-0002-8895-3206
- affiliation: 2
  name: Mattan S. Ben-Shachar
  orcid: 0000-0002-4287-4801
- affiliation: 3
  name: Indrajeet Patil
  orcid: 0000-0003-1995-6531
- affiliation: 4
  name: Brenton M. Wiernik
  orcid: 0000-0001-9560-6336
- affiliation: 5
  name: Etienne Bacher
  orcid: 0000-0002-9271-5075
- affiliation: 6
  name: Rémi Thériault
  orcid: 0000-0003-4315-6788
- affiliation: "7, 8"
  name: Dominique Makowski
  orcid: 0000-0001-5375-9967
affiliations:
- index: 1
  name:  University Medical Center Hamburg-Eppendorf, Germany
- index: 2
  name: Independent Researcher, Ramat Gan, Israel
- index: 3
  name: Carl Zeiss AG, Germany
- index: 4
  name: Independent Researcher, Tampa, FL, USA
- index: 5
  name: Luxembourg Institute of Socio-Economic Research (LISER), Luxembourg
- index: 6
  name: Department of Psychology, New York University, New York, NY, USA
- index: 7
  name: School of Psychology, University of Sussex, Brighton, UK
- index: 8
  name: Sussex Centre for Consciousness Science, University of Sussex, Brighton, UK
type: article
date: "`r Sys.Date()`"
bibliography: paper.bib
# abstract: |
#   The R ecosystem for statistical analysis is powerful but fragmented, requiring users to learn and combine numerous packages with inconsistent syntaxes. This presents a significant barrier to conducting a seamless and reproducible analytical workflow. The **easystats** project addresses this challenge by providing a unified and intuitive framework for statistical analysis in R. This ecosystem consists of a suite of interoperable packages, each designed for a specific stage of the statistical workflow. By offering a consistent design and syntax across packages, **easystats** simplifies the process of conducting, interpreting, and reporting statistical analyses. This lowers the cognitive load for users, making robust statistical methods more accessible to both novices and experts.
keywords: |
  R; easystats
acknowledgement: |
  *{easystats}* is part of the collaborative [*easystats*](https://github.com/easystats/easystats) ecosystem. Thus, we thank all [members of easystats](https://github.com/orgs/easystats/people), contributors, and users alike.
authorcontributions: |
  D.L. and D.M. drafted the paper; all authors contributed to both the writing of the paper and the conception of the software.
funding: |
  This research received no external funding.
conflictsofinterest: |
  The authors declare no conflict of interest.
output:
  rticles::joss_article
journal:
  pissn: 2475-9066
csl: apa.csl
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  dpi = 450,
  comment = "#>",
  fig.width = 6,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

# Summary

The **easystats** project is a collection of R packages that provides a unified and intuitive framework for data wrangling and statistical analysis. The ecosystem is built around a "tidyverse-like" philosophy of consistency, user-friendliness, and interoperability between packages. It aims to simplify the process of preparing, conducting, interpreting, and reporting statistical analyses by offering tools for a wide range of common tasks. These tasks include, for instance, data wrangling [`{datawizard}`, @patil_datawizard_2022], model assessment [`{performance}`, @Lüdecke2020performance], understanding and describing model parameters [`{parameters}`, @Lüdecke2020parameters], including Bayesian models [`{bayestestR}`, @Makowski2019], computation of effect sizes [`{effectsize}`, @Ben-Shachar2020], calculating and visualizing marginal effects [`{modelbased}`, @Makowski2025modelbased], and generating publication-ready figures [`{see}`, @ludecke_see_2021] or reports of statistical models [`{report}`, @report_remi_2023]. The **easystats** ecosystem is designed to be accessible to both novice and experienced R users, promoting reproducible research and a more seamless workflow from data exploration to result communication.

# Statement of Need

The R programming language is a powerful and dominant tool for statistical computing in academic research. However, its power is distributed across a vast and fragmented ecosystem of packages. For a researcher to conduct a comprehensive statistical analysis—from fitting a model to checking its assumptions, interpreting its parameters, and visualizing the results—they often need to learn and combine numerous packages, each with its own unique syntax, design principles, and output structures. This fragmentation presents a significant barrier to entry for new users and creates an inefficient and often confusing workflow, even for experienced analysts.

The **easystats** ecosystem addresses this challenge by providing a unified, consistent, and intuitive framework for data wrangling, data exploration, statistical summaries and modeling, visualization, and reporting in R. Its primary goal is to make robust and transparent statistical analysis more accessible to a broad range of users, from students to seasoned researchers. **easystats** achieves this through a suite of lightweight and interoperable packages, each designed to handle a specific stage of the statistical workflow. These packages share a consistent syntax and work seamlessly together, hence dramatically lowering the cognitive load required to conduct such analytical workflows.

Other software toolkits often serve different purposes. The `{tidyverse}` [@Wickham2019], for example, provides a world-class environment for data manipulation and general-purpose plotting but does not focus on the intricacies of statistical model interpretation and reporting. Specialist packages like `{lme4}` [@bates_fitting_2015] for mixed-effects models or `{marginaleffects}` [@arel-bundock_how_2024] for marginal effects and predictions are essential tools, but **easystats** serves as a complementary "meta-layer" that provides a single, easy-to-learn interface for interacting with the outputs from these and many other modeling packages. This allows researchers to focus on their scientific questions rather than on the technical idiosyncrasies of different software implementations. Hence, **easystats** fills a critical need in the R ecosystem for a user-friendly and cohesive suite of tools dedicated to the statistical modeling pipeline.

The modularity of the **easystats** packages enables developers to select and use only the necessary components. For example, `{insight}`, a dependency-free package for retrieving model information, is utilized by 45 other CRAN packages, and `{parameters}` is used by 22. In contrast, the `{easystats}` meta-package provides users with a cohesive experience, granting access to the entire ecosystem and its consistent design principles without needing to know the specific package of each function.

# A Harmonized and Integrated Workflow

A key design principle of the **easystats** ecosystem is the harmonization and integration of different packages into a simple, sequential workflow. The typical workflow for a statistical analysis using `{easystats}` starts with importing data and bringing the data into shape for the next step - fitting a model - and then sequentially using different functions to obtain a comprehensive understanding of the model.

Let's demonstrate this with an example, where the user starts by preparing some data and then fits a simple linear model:

```{r}
# we don't load each package individually,
# but rather the entire ecosystem
library(easystats)
data(coffee_data, package = "modelbased")

# dichotomize outcome variable
coffee_data$alertness <- categorize(coffee_data$alertness, lowest = 0)

# rename variable
coffee_data <- data_rename(coffee_data, select = c(treatment = "coffee"))

# fit model
model <- glm(
  alertness ~ treatment,
  data = coffee_data,
  family = binomial()
)
```

The `model` object can then be passed to functions from different **easystats** packages. For instance, the user can get a summary of the model parameters using the `{parameters}` package:

```{r}
model_parameters(model)
```

Then, the performance of the model can be assessed with the `{performance}` package:

```{r}
model_performance(
  model,
  metrics = c("AIC", "BIC", "R2", "PCP")
)
```

The results can be visualized using the `{see}` package, for example by plotting the model's predictions with `{modelbased}`:

```{r fig1, fig.cap = "Predicted probability of alertness by treatment group."}
predictions <- estimate_means(model, "treatment")
plot(predictions) + theme_modern(show.ticks = TRUE) # add nice theme
```

Finally, a full report of the analysis can be generated with the `{report}` package:

```{r}
report(model)
```

This seamless integration between packages allows users to move from model fitting to interpretation and reporting in a fluid and intuitive manner, without having to learn different syntaxes or data structures.

# Licensing and Availability

`{easystats}` is licensed under the MIT-License, with all source code stored at GitHub (<https://github.com/easystats/easystats>), and with a corresponding issue tracker for bug reporting and feature enhancements. In the spirit of honest and open science, we encourage requests, tips for fixes, feature updates, as well as general questions and concerns via direct interaction with contributors and developers.

# Acknowledgments

`{easystats}` is part of the collaborative [**easystats** ecosystem](https://easystats.github.io/easystats/). Thus, we thank all [members of easystats](https://github.com/orgs/easystats/people), contributors, and users alike.

# References
